---
title: "DropMap Natural Killer Cell Analysis"
name: Ali Ali
output: html_notebook
created:
---

### 1. IMAGE BATCH PROCESSING PREP - SPLIT TIFF IMAGE AND COPY CHANNEL 4 TO NEW DIRECTORY
```{r}
setwd("~/Desktop/test/input")
library(EBImage)
library(tiff)

#define the input directory and destination directory of the component of interest to be copied and used downstream
input_directory <- "~/Desktop/test/input/with_perforin"
destination_directory <- "~/Desktop/test/input/with_perforin/ilastic_input"

#create the destination directory if it doesn't exist
if (!dir.exists(destination_directory)) {
  dir.create(destination_directory)
}

#create a list of TIFF files in the directory
tiff_files <- list.files(input_directory, pattern = "\\.tif$", full.names = TRUE)

#process each TIFF file
for (input_filepath in tiff_files) {
  #load TIFF image
  img <- readImage(input_filepath)
  
  #get the number of channels in the image
  num_channels <- dim(img)[3]
  
  #get the base name of the input file (without the extension)
  input_basename <- tools::file_path_sans_ext(basename(input_filepath))
  
  #loop through each channel and save it as a separate image
  for (i in 1:num_channels) {
    # Extract the i-th channel
    channel_img <- img[,,i]
    
    #generate the output filename as "component_n" where n is the nth channel
    output_filename <- file.path(input_directory, paste0(input_basename, "_component_", i, ".tif"))
    
    #write the channel image to a new file
    writeImage(channel_img, output_filename)
    
    cat("Saved channel", i, "of", input_basename, "as", output_filename, "\n")
    
    #if the current channel is component_4, copy it to the destination folder
    if (i == 4) {
      destination_filename <- file.path(destination_directory, basename(output_filename))
      file.copy(output_filename, destination_filename)
      cat("Copied", output_filename, "to", destination_filename, "\n")
    }
  }
  
  cat("All channels of", input_basename, "have been saved.\n")
}

cat("All TIFF files have been processed.\n")

```

### 3. IMAGE BATCH PROCESSING PREP - RUN ILASTIC PIPELINE

### 2. IMAGE BATCH PROCESSING PREP - MIGRATE COMPONENT IMAGES OF PARENT TIFF TO FOLDER
```{r}
library(tools)

#define input directory
input_directory <- "~/Desktop/test/input"

#create a list of all files in the directory
files <- list.files(input_directory, full.names = TRUE)

#create a function to get the prefix of a file
get_prefix <- function(filename) {
  #extract the base name of the file (without the path)
  base_name <- basename(filename)
  
  #split the base name by underscores
  parts <- unlist(strsplit(base_name, "_"))
  
  #check if there are at least three parts (one underscore)
  if (length(parts) >= 3) {
    #join the parts before the third underscore to get the prefix
    prefix <- paste(parts[1:3], collapse = "_")
  } else {
    #if there are less than three parts, use the whole base name
    prefix <- base_name
  }
  
  return(prefix)
}

#loop through each file and move it to the corresponding folder
for (file in files) {
  #get the prefix of the file
  prefix <- get_prefix(file)
  
  #define the destination folder
  destination_folder <- file.path(input_directory, prefix)
  
  #create the destination folder if it doesn't exist
  if (!dir.exists(destination_folder)) {
    dir.create(destination_folder)
  }
  
  #define the destination file path
  destination_file <- file.path(destination_folder, basename(file))
  
  #move the file to the destination folder
  file.rename(file, destination_file)
  
  cat("Moved", file, "to", destination_file, "\n")
}

cat("All files have been migrated.\n")
```

### 4. INITIATE DATA ANALYSIS
```{r}
setwd("~/Desktop/test/input")
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
```

### IMPORT DATA
```{r}
#load cell count files and add time point column
data <- read.csv("objectcounts.csv")
```

### DATA PRE PROCESSING
```{r}
#split file name and extract condition and timepoint into new columns 
split_and_extract <- function(data, existing_col, new_col1, new_col2) {
  #create the presplit column by splitting the existing column on the first "_"
  data$presplit <- sapply(strsplit(data[[existing_col]], "_"), `[`, 1)
  
  #create the new condition and timepoint columns
  data <- data %>%
    mutate(!!new_col1 := str_extract(presplit, "[a-zA-Z ]+"),
           !!new_col2 := as.numeric(str_extract(presplit, "\\d+")))
  
  #remove the presplit column
  data$presplit <- NULL
  
  return(data)
}
data <- split_and_extract(data, "FileName_bubble_ilastik", "condition", "timepoint")

#trim data to include object number, counts, location, condition, and timepoint columns
trim_data <- function(data) {
  data <- data[, -c(1,3:12,21)]
  return(data)
}
data <- trim_data(data)

#rename cell count columns
rename_cols <- function(data) {
  data <- data %>% rename(
    deadkiller_count = Children_Identifydeadkiller_Count, 
    deadtarget_count = Children_Identifydeadtarget_Count, 
    alivekiller_count = Children_Identifyalivekiller_Count, 
    alivetarget_count = Children_Identifyalivetarget_Count,
    killer_count = Children_Identifykiller_Count,
    target_count = Children_Identifytarget_Count
  )
  return(data)
}
data <- rename_cols(data)

#add new column that contains info of ratio of target cells to killer cells in each bubble
ratio_col <- function(data) {
  data <- data %>% mutate(ratio_target_killer = target_count / killer_count) %>%
  mutate(ratio_target_killer = sprintf("%d:%d", target_count, killer_count)) %>%
  mutate(ratio_target_killer = factor(ratio_target_killer)
  )
  return(data)
}
data <- ratio_col(data)

#combine all data frames
combined_df <- data

#identify and filter artefacts
artefacts <- combined_df %>% filter(killer_count >= 4 | target_count >= 4)
combined_df <- combined_df %>% filter(killer_count < 4) %>% filter(target_count < 4)
```

### CREATE TARGET, KILLER, AND PAIRED DATA
```{r}
#target only bubbles - filter to include bubbles with targets and no killers
target_data <- combined_df  %>% 
  filter(target_count > 0 & killer_count == 0)

#killer only bubbles - filter to include bubbles with killers and no targets
killer_data <- combined_df  %>% 
  filter(killer_count > 0 & target_count == 0)

#target and killer bubbles - filter to include bubbles with both targets and killers
target_killer_data <- rbind(combined_df  %>% 
                              filter(deadtarget_count > 0 & deadkiller_count > 0), 
                            combined_df  %>% 
                              filter(deadtarget_count > 0 & alivekiller_count > 0), 
                            combined_df  %>% 
                              filter(alivetarget_count > 0 & deadkiller_count > 0), 
                            combined_df  %>% 
                              filter(alivetarget_count > 0 & alivekiller_count > 0)) %>% unique()
```

### BUBBLE AND CELL COUNT SUMMARY
```{r}
#killer only bubbles
killer_ctrl_summary1 <- rbind(
  #without perforin
  killer_data %>% 
    filter(condition == "without_perforin") %>%
    group_by(time_point) %>% 
    summarise(bubble_count = n(), condition, killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count), alivekiller_count = sum(alivekiller_count)),
  #with perforin
  killer_data %>% 
    filter(condition == "with_perforin") %>%
    group_by(time_point) %>% 
    summarise(bubble_count = n(), condition, killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count), alivekiller_count = sum(alivekiller_count))
  ) %>% 
  unique()

killer_ctrl_summary2 <- rbind(
  #without perforin
  killer_data %>% 
    filter(condition == "without_perforin") %>%
    group_by(ratio_target_killer) %>% 
    summarise(bubble_count = n(), condition, killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count), alivekiller_count = sum(alivekiller_count)),
  #with perforin
  killer_data %>% 
    filter(condition == "with_perforin") %>%
    group_by(ratio_target_killer) %>% 
    summarise(bubble_count = n(), condition, killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count), alivekiller_count = sum(alivekiller_count))
  ) %>% 
  unique()


#target only bubbles
target_ctrl_summary1 <- rbind(
  #without perforin
  target_data %>% 
    filter(condition == "without_perforin") %>%
    group_by(time_point) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), alivetarget_count = sum(alivetarget_count)),
  #with perforin
  target_data %>% 
    filter(condition == "with_perforin") %>%
    group_by(time_point) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), alivetarget_count = sum(alivetarget_count))
  ) %>% 
  unique()

target_ctrl_summary2 <- rbind(
  #without perforin
  target_data %>% 
    filter(condition == "without_perforin") %>%
    group_by(ratio_target_killer) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), alivetarget_count = sum(alivetarget_count)),
  #with perforin
  target_data %>% 
    filter(condition == "with_perforin") %>%
    group_by(ratio_target_killer) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), alivetarget_count = sum(alivetarget_count))
  ) %>% 
  unique()



#paired bubbles
exp_summary1 <- rbind(
  #without perforin
  target_killer_data %>% 
    filter(condition == "without_perforin") %>%
    group_by(time_point) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count)),
  #with perforin
  target_killer_data %>% 
    filter(condition == "with_perforin") %>%
    group_by(time_point) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count))
  ) %>% 
  unique()

exp_summary2 <- rbind(
  #without perforin
  target_killer_data %>% 
    filter(condition == "without_perforin") %>%
    group_by(ratio_target_killer) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count)),
  #with perforin
  target_killer_data %>% 
    filter(condition == "with_perforin") %>%
    group_by(ratio_target_killer) %>% 
    summarise(bubble_count = n(), condition, target_count = sum(target_count), deadtarget_count = sum(deadtarget_count), killer_count = sum(killer_count), deadkiller_count = sum(deadkiller_count))
  ) %>% 
  unique()
```

### DISTRIBUTION PLOTS
```{r}
#killer distribution in droplets
ggplot(data.frame(combined_df %>% group_by(killer_count) %>% summarise(count = n()) %>% mutate(pct = (count/sum(count))*100)), aes(x = killer_count, y = pct)) +
   geom_bar(stat = "identity", width = 0.5, fill = "#CC79A7") +
   labs(title = "NK distribution in droplets", x = "Killer Count", y = "% droplets with cells") +
   ylim(0, 100) +
   theme_classic()

#target distribution in droplets
ggplot(data.frame(combined_df  %>% group_by(target_count) %>% summarise(count = n()) %>% mutate(pct = (count/sum(count))*100)), aes(x = target_count, y = pct)) +
   geom_bar(stat = "identity", width = 0.5, fill = "#56B4E9") +
   labs(title = "YAC distribution in droplets", x = "Target Count", y = "% droplets with cells") +
   ylim(0, 100) +
   theme_classic()

#cell pairing in droplets
ggplot(data.frame(
  bubble = c("paired", "unpaired killers", "unpaired targets"), 
  pct = c(
    (nrow(target_killer_data) / (nrow(target_killer_data) + nrow(killer_data) + nrow(target_data))) * 100, 
    (nrow(killer_data) / (nrow(target_killer_data) + nrow(killer_data) + nrow(target_data))) * 100, 
    (nrow(target_data) / (nrow(target_killer_data) + nrow(killer_data) + nrow(target_data))) * 100)), 
  aes(x = bubble, y = pct)) +
   geom_bar(stat = "identity", width = 0.5, fill = c("#E69F00" , "#CC79A7", "#56B4E9")) +
   labs(title = "Cell Pairing in Droplets", x = "droplet type", y = "% droplets with cells") +
   ylim(0, 100) +
   theme_classic()

#T:K ratio distribution
ggplot(data.frame(combined_df %>%group_by(ratio_target_killer) %>% summarise(count = n()) %>% mutate(pct = (count/sum(count))*100)), aes(x = ratio_target_killer, y = pct)) +
   geom_bar(stat = "identity", fill = "#E69F00") +
   labs(title = "T:K Ratio Distribution", x = "T:K Ratios in Bubbles", y = "% droplets") +
   ylim(0, 100) +
   theme_classic()

#killer and target counts in paired droplets
ggplot(data.frame(cell = c("killer", "target"), sum = c(sum(target_killer_data$killer_count), sum(target_killer_data$target_count))) %>% 
                    mutate(pct = (sum/(sum(target_killer_data$killer_count) + sum(target_killer_data$target_count)) * 100)), 
       aes(x = cell, y = pct)) +
   geom_bar(stat = "identity", width = 0.5, color = "#E69F00", fill = c("#CC79A7", "#56B4E9")) +
   labs(title = "Killer and Target Counts in Paired Droplets", x = " ", y = "% cells in droplets") +
   ylim(0, 100) +
   theme_classic()
```

### CYTOTOXCITY PLOTS
```{r}
cytotox1 <- rbind(
  target_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "target_only") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count", "target_count", "killer_count")) %>% unique(),
  
   target_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "target_only") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count", "target_count", "killer_count")) %>% unique(),
  
  killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "killer_only") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count", "target_count", "killer_count")) %>% unique(),
  
  killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "killer_only") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count", "target_count", "killer_count")) %>% unique(),
  
  target_killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "paired") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count", "target_count", "killer_count")) %>% unique(),
  
    target_killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "paired") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count", "target_count", "killer_count")) %>% unique(),
) %>% mutate(pct = (death/total_cell_count) * 100)

ggplot(cytotox1, aes(x = time_point, y = pct, color = bubble_type)) + 
 geom_line() + 
 labs(title = "Cytotoxcity Profile", x = "time (hr)", y = "% caspase postive cells") +
 facet_wrap(vars(condition)) +
 scale_color_manual(values = c("killer_only" = "#56B4E9", "paired" = "#E69F00", "target_only" = "#CC79A7"),
                    labels = c("killer_only" = "killer only", "target_only" = "target only")) +
 ylim(0, 100) +
 theme_classic()  +
 theme(legend.position = "bottom")

#ggsave(filename = "cytotox1.png", width = 2000, height = 1000, units = "px", plot = cytotox1)



cytotox2 <- rbind(
  target_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "target_only") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
   target_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "target_only") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
  killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "killer_only") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
  killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "killer_only") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count, deadkiller_count), total_cell_count = sum(target_count, killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
  target_killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "paired_targetdeath") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count), total_cell_count = sum(target_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
    target_killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "paired_targetdeath") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadtarget_count), total_cell_count = sum(target_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
    target_killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "paired_killerdeath") %>% filter(condition == "without_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadkiller_count), total_cell_count = sum(killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique(),
  
    target_killer_data[, c("condition", "time_point", "deadtarget_count", "deadkiller_count", "target_count", "killer_count")] %>% mutate(bubble_type = "paired_killerdeath") %>% filter(condition == "with_perforin") %>% group_by(time_point) %>% mutate(death = sum(deadkiller_count), total_cell_count = sum(killer_count)) %>% select(-c("deadtarget_count", "deadkiller_count")) %>% unique()
) %>% mutate(pct = (death/total_cell_count) * 100)

ggplot(cytotox2, aes(x = time_point, y = pct, color = bubble_type)) + 
 geom_line() + 
 labs(title = "Cytotoxcity Profile", x = "time (hr)", y = "% caspase postive cells") +
 facet_wrap(vars(condition)) +
 scale_color_manual(values = c("killer_only" = "#56B4E9", "paired_killerdeath" = "#E69F00", "paired_targetdeath" = "#F0E442", "target_only" = "#CC79A7"), 
                    labels = c("killer_only" = "killer only (killer death)", "paired_killerdeath" =  "paired (killer death)", "paired_targetdeath" = "paired (target death)", "target_only" = "target only (target death)")) +
 ylim(0, 100) +
 theme_classic() +
 theme(legend.position = "bottom")

#ggsave(filename = "cytotox2.png", width = 2000, height = 1000, units = "px", plot = cytotox2)
```
